Experiment 2 (of 4)

### 2016-08-10 experiments/e2/writeup/e2-b-knitr.Rnw
Switched to knitr latex Rnw

---
title: "Experiment 2 (of 4)"
author: "Matt Green"
output: 
  pdf_document:
    toc: true
---

```{r loadKnitr, echo=FALSE}
suppressPackageStartupMessages(library(knitr))
```

```{r loadLibs, echo=FALSE, message=FALSE}
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
```

```{r get_the_data, echo=FALSE}

@@

```

Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.

```{r plotDists, fig.width=5, fig.height=5, message=FALSE, echo=FALSE}
# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
load('data_raw.Rda') # dat
dat_transforms <- dat
# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"
# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"
# create extra rows for dfs with removed (outlier) data points
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"
# bring the dfs back into main df
dat_transforms <- rbind(dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("RT>1&RT<59999", "RT>1&RT<30001"))
# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_log", "RT_raw")] <- NA
# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  }# end of create normally distributed equivalents for the data transforms.
temp <- reshape2::melt(dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_log", "RT_raw"),
             variable.name="transformation",
             value.name="time")
ggplot(temp, aes(time, colour=RTtype)) + geom_freqpoly() + facet_wrap(transformation~dfSubset, scales='free') + scale_color_manual(values=c("blue","red")) + theme_bw() + theme(aspect.ratio=1)
```

# Now working with a data set without impossible trials.

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r boxPlots, fig.width=7, fig.height=3, echo=FALSE}
load('data_processed.Rda')
# allow reference to transformation
subdata = reshape2::melt(dd,
               measure.vars=c("RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
  xlab("") + ylab("")
```

Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.

```{r fastSlowRT, fig.width=7, fig.height=3, message=FALSE, echo=FALSE}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), 
					 summarise, effect="subject", emean=mean(dd$RT_log), lmean=mean(RT_log),
					 ldiff=mean(RT_log)-mean(dd$RT_log))
itms=ddply(dd, .(level=Item), 
					 summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
					 ldiff=mean(RT_log)-mean(dd$RT_log))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col='black'), 
              																	superpose.line=list(col='black'))),
              xlab=NULL,
              ylab=NULL,
              main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=NULL, 
              									 labels=list("subject number", "item identifier"), cex=1 ),
                          y=list(at=    c(0,7,unique(si$emean),8,9),
                          			 labels=c("0","7","Mean Log RT","8","9"))),
              panel=function(x,y,type,subscripts,...){
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, subscripts=subscripts, code=3, length=0, ...)
                panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
              }) # end of my1
my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from\nmean Log RT",
              ylim=extendrange(si$ldiff,f=0.10)
							) # creates a second y-axis scale
print(doubleYScale(my1, my2, 
									 add.axis=TRUE, 
									 add.ylab2=TRUE, 
									 use.style=FALSE))
```

# Plot main effects and interactions

Main effects

Plot main effects on several transformations

```{r plotMainsa, fig.width=6, fig.height=6, warning=FALSE, echo=FALSE}
# plot main effects on observed data
# allow reference to transformation
temp <- reshape2::melt(dd,
             id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
             variable.name="transformation",
             value.name="score")
# allow reference to effect
temp2 <- reshape2::melt(temp, 
              id.vars=c("transformation", "score"),
              measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
              variable.name="effect",
              value.name="level")
# do the plot
ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
  facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
  stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
  stat_summary(fun.y=mean, geom="line") +
  theme(aspect.ratio=1, axis.text.y=element_blank())
```

Main effects

```{r mainB, fig.width=6, fig.height=3, echo=FALSE}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
# plot main effects
ggplot(rts, aes(y=RT_RecSqt, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r mainD, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Number")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
ggplot(rts, 
			 aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, 
			 		ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
  scale_color_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
```

2 way interactions not involving items.

```{r 2way, fig.width=6, fig.height=3, echo=FALSE}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Number")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Number",4)
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Number:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Number:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, sd, se, ci))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
# plot 2 way interactions
ggplot(rts, aes(y=RT_RecSqt, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

>>>>

\documentclass[a4paper]{article}
\usepackage{geometry}
\usepackage{placeins}
\begin{document}

<<loadKnitr, echo=FALSE, message=FALSE>>=
library(knitr)
opts_chunk$set(cache=FALSE, autodep = TRUE)
@

<<defineFonts>>=
# .define.fonts <- function () {
#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
# }
# .define.fonts()
@

<<loadFonts>>=
library(extrafont)
loadfonts()
@

<<loadLibs, echo=FALSE, message=FALSE>>=
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
@

\section{Get the data}

<<getTheData, echo=FALSE>>=

@@

@

\section{Plots}

<<prepPlotData, echo=FALSE>>=
# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
load('data_raw.Rda') # dat
dat_transforms <- dat
# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"
# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"
# create extra rows for dfs with removed (outlier) data points
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"
# bring the dfs back into main df
dat_transforms <- rbind(dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("RT>1&RT<59999", "RT>1&RT<30001"))
# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_log", "RT_raw")] <- NA
# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  }# end of create normally distributed equivalents for the data transforms.
temp <- reshape2::melt(dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_log", "RT_raw"),
             variable.name="transformation",
             value.name="time")
@

<<plotDists, fig.width=5, fig.height=5, fig.cap='Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions', fig.pos='hbtp', fig.align='right', message=FALSE, warning=FALSE, echo=FALSE>>=
ggplot(temp, aes(time, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(transformation~dfSubset, scales='free') + 
  scale_color_manual(values=c("blue","red")) + 
  theme_bw() + 
  theme(aspect.ratio=1, text=element_text(family='CMU Serif', size=12, face='italic'))
@

<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE>>=
p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
ggtitle("Fuel Efficiency of 32 Cars") +
xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
theme_bw() +
theme(text=element_text(family="CMU Serif", size=14, face='plain'))
p
ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
@

<<boxPlots, fig.width=7, fig.height=3, fig.cap='Show how transformations of RT affect distribution of times, and how they affect which times are outliers.', echo=FALSE>>=
load('data_processed.Rda')
# allow reference to transformation
subdata = reshape2::melt(dd,
                         measure.vars=c("RT_log", "RT_raw"),
                         variable.name="transformation",
                         value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
  xlab("") + ylab("")
  @
  <<fastSlowRT, fig.width=7, fig.height=3, fig.cap='Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.', message=FALSE, echo=FALSE>>=
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), 
summarise, effect="subject", emean=mean(dd$RT_log), lmean=mean(RT_log),
ldiff=mean(RT_log)-mean(dd$RT_log))
itms=ddply(dd, .(level=Item), 
summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
ldiff=mean(RT_log)-mean(dd$RT_log))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
my1 <- xyplot(data=si, lmean~level | effect, 
aspect=1,
layout=c(2,1),
ylim = extendrange(si$lmean, f = 0.10),
par.settings=list(trellis.par.set(superpose.symbol = list(col='black'), 
superpose.line=list(col='black'))),
xlab=NULL,
ylab=NULL,
main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
mean=unique(si$emean),
scales=list(x=list(relation="free", at=NULL, 
labels=list("subject number", "item identifier"), cex=1 ),
y=list(at=    c(0,7,unique(si$emean),8,9),
labels=c("0","7","Mean Log RT","8","9"))),
panel=function(x,y,type,subscripts,...){
panel.abline(h=unique(si$emean), col="lightgrey",lty=1,...)
panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, subscripts=subscripts, code=3, length=0, ...)
panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
}) # end of my1
my2 <- xyplot(data=si, ldiff~level | effect, 
aspect=1, type='n',
ylab="Deviation from\nmean Log RT",
ylim=extendrange(si$ldiff,f=0.10)
) # creates a second y-axis scale
print(doubleYScale(my1, my2, 
add.axis=TRUE, 
add.ylab2=TRUE, 
use.style=FALSE))
@

<<plotMainsa, fig.width=6, fig.height=6, fig.cap='Plot main effects on several transformations', message=FALSE, echo=FALSE, warning=FALSE>>=
# plot main effects on observed data
# allow reference to transformation
temp <- reshape2::melt(dd,
id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
variable.name="transformation",
value.name="score")
# allow reference to effect
temp2 <- reshape2::melt(temp,
id.vars=c("transformation", "score"),
measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
variable.name="effect",
value.name="level")
# do the plot
ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
stat_summary(fun.y=mean, geom="line") +
theme(aspect.ratio=1, axis.text.y=element_blank())
@

<<mainB, fig.width=6, fig.height=3, fig.cap='Just the main effects', message=FALSE, echo=FALSE>>=
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
# plot main effects
ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) +
geom_line() +
geom_errorbar(width=.1) +
geom_point(cex=3) +
facet_grid(~Effect, scale="free_x") +
theme_bw() +
theme(aspect.ratio=1)
@

<<mainD, fig.width=6, fig.height=3.5, fig.cap='Main effects over item ratios', message=FALSE, echo=FALSE>>=
nrts <- cbind(Effect="Number",
xlev=factor(c(1,2)), 
myshape='a',
summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Number")))
names(nrts)[5]<-"Levels"

orts <- cbind(Effect="Order",
xlev=factor(c(1,2)), 
myshape='b',
summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")))
names(orts)[5]<-"Levels"

qrts <- cbind(Effect="Quantity",
xlev=factor(c(1,2)), 
myshape='d',
summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")))
names(qrts)[5]<-"Levels"

vrts <- cbind(Effect="Vagueness",
xlev=factor(c(1,2)),
myshape='e',
summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")))
names(vrts)[5]<-"Levels"

rts=rbind(nrts,orts,qrts,vrts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
#rts$myshape=rep(c(21,22,23,24), each=8)

p1=
ggplot(data=rts, aes(x=item_mean_ratio, y=RT_log, group=Levels)) +
geom_errorbar(data=droplevels(subset(rts, Effect=='Number')),width=.01, aes(ymin=mins, ymax=maxs)) +
geom_line(data=droplevels(subset(rts, Effect=='Number'))) +
geom_point(data=droplevels(subset(rts, Effect=='Number')),
aes(shape=Levels, fill=Levels), size=3 ) +
geom_errorbar(data=droplevels(subset(rts, Effect=='Order')), width=.01, aes(ymin=mins, ymax=maxs)) +
geom_line(data=droplevels(subset(rts, Effect=='Order'))) +
geom_point(data=droplevels(subset(rts, Effect=='Order')),
aes(shape=Levels, fill=Levels), size=3  ) +
geom_errorbar(data=droplevels(subset(rts, Effect=='Quantity')),width=.01, aes(ymin=mins, ymax=maxs)) +
geom_line(data=droplevels(subset(rts, Effect=='Quantity'))) +
geom_point(data=droplevels(subset(rts, Effect=='Quantity')),
aes(shape=Levels, fill=Levels), size=3  ) +
geom_errorbar(data=droplevels(subset(rts, Effect=='Vagueness')),width=.01, aes(ymin=mins, ymax=maxs)) +
geom_line(data=droplevels(subset(rts, Effect=='Vagueness'))) +
geom_point(data=droplevels(subset(rts, Effect=='Vagueness')),
aes(shape=Levels, fill=Levels), size=3  ) +
facet_grid(~Effect) +
scale_shape_manual(values = c(
"Numeric"=21,
"Verbal"=21,
"LtoR"=22,
"RtoL"=22,
"Small"=23,
"Large"=23,
"Crisp"=24,
"Vague"=24),
breaks=c('Numeric','Verbal','LtoR','RtoL','Small','Large','Crisp','Vague'))  +
scale_fill_manual(values = c(
"Numeric"="black",
"Verbal"="white",
"LtoR"="black",
"RtoL"="white",
"Small"="black",
"Large"="white",
"Crisp"="black",
"Vague"="white"),
breaks=c('Numeric','Verbal','LtoR','RtoL','Small','Large','Crisp','Vague')) +
scale_x_continuous(minor_breaks=NULL,
breaks=c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
labels=abbreviate(c(0.5125000, 0.6876471, 0.7691558, 0.8166667))) +
theme(aspect.ratio=1,
legend.position='top',
axis.text.x=element_text(size=7),
legend.key=element_rect(fill = 'white'),
panel.background=element_rect(fill = 'grey96'))
p1
@

<<22way, fig.width=6, fig.height=3, fig.cap='2-way interactions over items', message=FALSE, echo=FALSE>>=
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Number"
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Number"
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Number"
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"
rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
rts$F3 <- factor(paste(rts$F1,rts$F2))
rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) +
facet_grid(E2~E1, drop=F) +
geom_point(cex=3) +
geom_line() +
theme_bw() +
theme(aspect.ratio=1, panel.grid=element_blank()) +
scale_color_manual(name="colour",values=cbbPalette) +
scale_shape_discrete(name="shape") +
scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
@

<<plotInt1, fig.width=7, fig.height=3.5, fig.cap='Plot vagueness by number interaction over items', echo=FALSE>>=
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
group=condition, shape=condition, fill=condition)) +
geom_line(position=dodge) +
geom_errorbar(width=.2, position=dodge) +
geom_point(size=4, position=dodge) +
scale_shape_manual("",values = c(22, 22, 21, 21)) +
scale_fill_manual("",values=c("black","white","black","white")) +
ggtitle("Response time") +
ylab("RT_RecSqt") +
xlab("") +
theme_bw() +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
legend.key = element_blank(), aspect.ratio=1,
axis.text.x = element_text(angle = 45, hjust = 1)) +
facet_grid(~Number)
@
